---
title: "Group 4 Project"
author: "Laura Larregui, Angela Garcia, John Christman"
date: "11/18/2019"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r readWines}
readURL <- function(inputURL)  #Begin function named readURL that takes a URL
{
  csvFile <- read.csv(url(inputURL), sep = ';')  #assign the results of the URL call as a csv file to a dataframe named csvFile. Added sep = ';' to seperate the data into columns
  return(csvFile)  # return the dataframe
}

redWine <- readURL("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv")
whiteWine <- readURL("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv")

str(redWine)
str(whiteWine)
summary(redWine)
summary(whiteWine)


```
```{r cleanWine}
#THe datasets only have one column of data.  The column names are separated by periods the data by semi-colons
#1. Create columns 
#2.  separate the data into the columns
#3.  Verify no NAs
redWine <- na.omit(redWine)
whiteWine<-na.omit (whiteWine)

```

<<<<<<< HEAD
```{r visualizeWine }
#1.  Create visulaizations for the data
=======
```{r Standard Deviations}
#Quality Standard Deviations
sd(redWine$quality)
sd(whiteWine$quality)

#Red wine Standard Deviations
sd(redWine$quality)
sd(redWine$alcohol)
sd(redWine$residual.sugar)
sd(redWine$pH)

#White wine standard deviations
sd(whiteWine$quality)
sd(whiteWine$alcohol)
sd(whiteWine$residual.sugar)
sd(whiteWine$pH)

```

```{r visualizeWine echo=FALSE}
#1.  Create visualizations for the data
>>>>>>> bfdd2188e5a5ba218bd22cdd1fe4e61166fd2420
#heat maps, histograms and scatter plots?

#Heatmaps
red_cor <- cor(redWine)
white_cor <- cor(whiteWine)
col<- colorRampPalette(c("blue", "white", "red"))(20)
heatmap(x = red_cor, col = col, symm = TRUE)
heatmap(x = white_cor, col = col, symm = TRUE)

#Histograms

#Histograms
hist(redWine$quality, main = "Red Wine Distribution", xlab = "Quality with Mean = 5.636 and SD = 0.8076", col ="red4")
hist(whiteWine$quality, main = "White Wine Distribution", xlab = "Quality with Mean = 5.878 and SD = 0.8856", col = "lemonchiffon")

library(grid)
library(gridExtra)
library (ggplot2)

h1 <- ggplot(aes(density), data = redWine) + geom_histogram(bins = 30,fill= "tomato3",color="white")
h1 <- h1 + ggtitle("Density Distribution") +theme(axis.title.x = element_blank())

h2 <- ggplot(aes(alcohol), data = redWine) + geom_histogram(bins = 30, fill= "tomato3",color="white")
h2 <- h2 + ggtitle("Alcohol Distribution") +theme(axis.title.x = element_blank())

h3 <- ggplot(aes(residual.sugar), data = redWine) + geom_histogram(bins = 7, fill= "tomato3",color="white")
h3 <- h3 + ggtitle("Sugar Distribution") + theme(axis.title.x = element_blank())

h4 <- ggplot(aes(pH), data = redWine) + geom_histogram(bins = 7, fill= "tomato3",color="white")
h4 <- h4 + ggtitle("pH Distribution")+theme(axis.title.x = element_blank())

grid.arrange(h1,h2,h3,h4,ncol=2)

g1 <- ggplot(aes(density), data = whiteWine) + geom_histogram(bins = 25,fill= "lightgoldenrod2",color="white")
g1 <- g1 +  ggtitle("Density Distribution") + theme(axis.title.x = element_blank())

g2 <- ggplot(aes(alcohol), data = whiteWine) + geom_histogram(bins = 20, fill= "lightgoldenrod2",color="white")
g2 <- g2 + ggtitle("Alcohol Distribution") + theme(axis.title.x = element_blank())

g3 <- ggplot(aes(residual.sugar), data = whiteWine) + geom_histogram(bins = 7, fill= "lightgoldenrod2",color="white")
g3 <- g3 + ggtitle("Sugar Distribution") + theme(axis.title.x = element_blank())

g4 <- ggplot(aes(pH), data = whiteWine) + geom_histogram(bins = 7, fill= "lightgoldenrod2",color="white")
g4 <- g4 + ggtitle("pH Distribution") + theme(axis.title.x = element_blank())

grid.arrange(g1,g2,g3,g4,ncol=2)

```

```{r correlationWine}
#1. Create the correlation matrix

#Red Wine Correlation Matrix

#install.packages("corrplot")
library(corrplot)
red_cor <- cor(redWine)
round(red_cor, 2)
corrplot(red_cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
#Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients.

#Correlation matrix with numbers
corrplot(red_cor, method = 'number', type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

#White Wine Correlation Matrix

#install.packages("corrplot")
library(corrplot)
white_cor <- cor(whiteWine)
round(white_cor, 2)
corrplot(white_cor, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
#Correlation matrix with numbers
corrplot(white_cor, method = 'number',type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

#Quality correlations
dfred_cor<-data.frame(red_cor)
dfwhite_cor<-data.frame(white_cor)
names<-row.names(dfred_cor)

QualityCor<-round(cbind(dfred_cor$quality,dfwhite_cor$quality),digits=4)
colnames(QualityCor)<-c("Red Quality", "White Quality")
row.names(QualityCor)<-names
View(QualityCor)

##Reference: http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
```

```{r learingWine}
#Machine learning techniques to see if we can train the system to pick a good wine
```


